---
title: "Iris clustering"
author: "Juan M. Carrillo"
date: "September 12, 2018"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objective

The main objective is to create an interactive `shiny` application.

I decide to let the user cluster a dataset into different groups according to the characteristics of its features.

The dataset to cluster will be the ubiquitous `iris` dataset since probably the user is familiarized with it.

## Inputs

The user will be presented with four different inputs:
 
 
On the side panel:

- Number of clusters: From 2 to 10 the user will decide how many groups the data set will be clustered in.
- X axis: The plot generated will use the x-axis chosen by the user.
- Y axis: The plot will use the y-axis selected.
 
 
On the main panel:

- By dragging the pointer on the generated plot, the user will be able to select any point.

## Output

Two will be the elements shown as a output on the main panel:

- A plot with the axis selected by the user and each case represented and colored by its cluster and the centroid of the cluster.
- A table with information from the `iris` dataset of the selected points.


## Main plot

Each case is assigned to a cluster using the `kmeans` function from the `stat` package.

The user should see a plot similar to the following one:

```{r plot, echo = F}
library (ggplot2)
set.seed (13920)

irisWoSpecies <- iris [, -5] # Remove the `Species` variable
clustResult <- kmeans (x = irisWoSpecies, centers = 3)

clustIris <- cbind (iris, cluster = clustResult$cluster)
centroids <- as.data.frame (clustResult$centers)

for (i in 1:nrow (centroids)) {
    centroids$cluster[i] <- i 
}

ggplot (data = clustIris, 
        aes (x = clustIris [, "Petal.Length"],
             y = clustIris [, "Petal.Width"])) +
    geom_point (aes (color = as.factor (cluster)) ) +
    geom_point (data = centroids,
                aes (x = centroids [, "Petal.Length"],
                     y = centroids [, "Petal.Width"], color = factor (cluster)),
                shape = 4, size = 10) +
    theme (plot.title = element_text (size = 12)) +
    labs (title = "Clustering of the Iris data set",
          x = "Petal.Length",
          y = "Petal.Width",
          color = "Clusters")
```

